Structural 

### 2018-11-26 r/SEM.Rmd
11/26

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(here)
library(tidyverse)
library(lavaan)
library(semPlot)
```

```{r,message=FALSE,warning=FALSE, echo=FALSE}
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=FALSE) %>% select(-X)
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=FALSE) %>% select(-X)

# recode phase and species with dummy variables
# data <- data %>% mutate(spec.dum= 
#                           ifelse(Species.Code=="stop", 1, 
#                           (ifelse(Species.Code=="qup", 2,
#                           (ifelse(Species.Code=="rbp",3,NA)))))) %>%
#   mutate(phase.dum=ifelse(Phase=="i",1,2))

@@

```{r}
# testing partial SEMs

sem_df <- sum_id %>% 
  filter(island!="Barbuda") %>%
  select(fr, g_frac, species_code, phase, scar_bm, scar_den, carn30_bm, tot_bm, spear_daily, ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, int_scar, length_cm, lc_cover) %>% 
  mutate(phase_dum=ifelse(phase=="i",1,2),
         fr_sc = fr/100,
         scar_bm_sc = scar_bm/100,
         rugosity_sc = rugosity*10) %>%
  na.omit()
vet_df <- sem_df %>% filter(species_code=="qup")
vir_df <- sem_df %>% filter(species_code=="stop")

@@

fit <- sem(sem3, data=vet_df, missing="ML", std.lv=T, orthogonal=T)
varTable(fit)
summary(fit, standardized=T, fit.measures=T, rsq=T)
semPaths(fit)
fitMeasures(fit, c("cfi","rmsea","srmr"))
```

>>>>

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(here)
library(tidyverse)
library(lavaan)
library(semPlot)
library(piecewiseSEM)
```

```{r,message=FALSE,warning=FALSE, echo=FALSE}
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=FALSE) %>% select(-X)
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=FALSE) %>% select(-X)
sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=FALSE) %>% select(-X)
```

# Full model layout

@@

# ID level SEMs

```{r}
# testing partial SEMs with ID - level dataset

sem_df <- sum_id %>% 
  # filter(island!="Barbuda") %>%
  filter(species_code != "rbp") %>%
  select(fr, g_frac, species_code, phase, scar_bm, scar_den, carn30_bm, tot_bm, spear_daily, ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, int_scar, length_cm, lc_cover) %>% 
  mutate(phase_dum=ifelse(phase=="i",0,1),
         species_dum = ifelse(species_code=="stop",0,1),
         fr_sc = fr/100,
         scar_bm_sc = scar_bm/100,
         rugosity_sc = rugosity*10) %>%
  na.omit()
vet_df <- sem_df %>% filter(species_code=="qup")
vir_df <- sem_df %>% filter(species_code=="stop")

@@

fit <- sem(sem1, data=sem_df, missing="ML", std.lv=T, orthogonal=T)
varTable(fit)
summary(fit, standardized=T, fit.measures=T, rsq=T)
semPaths(fit)
fitMeasures(fit, c("cfi","rmsea","srmr"))
```

# Site-level SEM

```{r}
sem_df <- sum_ssp %>% 
  select(fr_mean, species_code, phase, scar_bm, scar_den, carn30_bm, tot_bm, spear_daily, ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, int_scar_mean, length_m, lc_cover) %>% 
  filter(fr_mean > 0,
         species_code!="rbp") %>%
  mutate(phase_dum = ifelse(phase=="i",0,1),
         species_dum = ifelse(species_code=="stop",0,1),
         fr_sc = fr_mean/100,
         scar_bm_sc = scar_bm/100,
         rugosity_sc = rugosity*10) %>%
  na.omit()
vet_df <- sem_df %>% filter(species_code=="qup")
vir_df <- sem_df %>% filter(species_code=="stop")

sem1 <- '
  # latent variable indicators
      risk =~ scar_bm_sc + scar_den + rugosity_sc + spear_daily
      nutrition =~ ma_cover + ta_cover
  # regressions
      fr_sc ~ species_dum + length_m + phase_dum + nutrition + risk
      rugosity_sc ~ lc_cover
      lc_cover ~ ma_cover + ta_cover
'

sem2 <- '
  # latent variable indicators
      risk =~ scar_bm_sc + scar_den
      nutrition =~ ma_cover + ta_cover
  # regressions
      fr_sc ~ species_dum + length_m + phase_dum + nutrition + risk
'

fit <- sem(sem2, data=sem_df, missing="ML", std.lv=T)
varTable(fit)
summary(fit, standardized=T, fit.measures=T)
semPaths(fit, 'std', residuals = F)
fitMeasures(fit, c("cfi","rmsea","srmr"))

# Piecewise SEM

```{r}
sum_id_nar <- sum_id %>% drop_na()
psem1 <- psem(
  lm(fr ~ phase + species_code + scar_bm + ta_cover, sum_id_nar),
  lm(ta_cover ~ scar_bm, sum_id_nar))
summary(psem1)
```
